---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# cecmidas

<!-- badges: start -->
<!-- badges: end -->

cecmidas is an R package designed for interaction and uploading to the California Energy Commission's (CEC) Market Informed Demand Automation Server (MIDAS). MIDAS provides access to real-time and historical electricity rates for the largest electricial utilities in California. Documentation for MIDAS is available at: [MIDAS-Documentation](https://gitlab.com/CEC-MIDAS/midas-documentation). Reading the MIDAS documentation is highly recommended as it makes using this package much easier.

cecmidas can both GET and POST to MIDAS, but most users will only have GET access. POST access is limited to the California utilities that maintain their rates in MIDAS. The package includes some helper functions for those utilities to help prep and upload their rate data.

This package is under active development. One of the next priorities is to add a way for handling secrets so that you don't expose passwords in your code. If you have issues or feature requests, please use [GitHub issues](https://github.com/stefwayland/cecmidas/issues) to provide feedback.

## Installation

<!-- You can install the released version of cecmidas from [CRAN](https://CRAN.R-project.org) with: -->

<!-- ``` r -->
<!-- install.packages("cecmidas") -->
<!-- ``` -->

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("stefwayland/cecmidas")
```
## Getting Started

Please start by referring to the [MIDAS Documentation](https://gitlab.com/CEC-MIDAS/midas-documentation).

Using MIDAS requires that you first register with the server. You only need to do this once. MIDAS will send a conifrmation email to the address you give it. Click through the confirmation link and you will be good to go! (Unless you are a utility in California who needs upload access, then see the [MIDAS documentation](https://gitlab.com/CEC-MIDAS/midas-documentation) for next steps.)

cecmidas uses [R6](https://r6.r-lib.org/) objects, so the code here looks a little different than standard R code. The first step to using cecmidas is to create a `midas` object (you can name it whatever you want) which holds your login information and connection token. This makes it easier to use once you get started because you don't have to provide that information to every function or store it in an environment. To do this yourself, just replace the username etc., in the example with your own.

```{r register, eval = FALSE}
library(cecmidas)
# Create a MIDAS object
midas <- MIDAS$new(username = "alovelace",
                   password = "Differenceengineftw!",
                   email = "alovelace@example.com",
                   fullname = "Ada Lovelace",
                   organization = "First Computing Inc")
# Register
midas$register()
```

Once you've registration process the fist time, you will not have to use the `register()` function again. Just create a midas object and get on with it. For example, to get the list of the rates available and then look up the current price for one of those rates:

```{r example, eval = FALSE}
library(cecmidas)
# Create a MIDAS object (only do this one per session)
midas <- MIDAS$new(username = "alovelace",
                   password = "Differenceengineftw!")

# Get list of Rate Identification Numbers (RINs) and put them in a data.frame
rins <- midas$rins()
rins

# Get real-time data for the rate with RIN "USCA-TSTS-TTOU-TEST"
prices <- midas$value(rin = "USCA-TSTS-TTOU-TEST", query_type = "realtime")
prices$ValueInformation
```
